---
title: "prior"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Find appropriate b. Find appropriate b and alpha for training.
2. Find appropriate prior for topics.

```{r}
library(tidyverse)
```

### corpus memberships

Find appropriate b (expectation of a) for generating data. We want b large enough to generate relative balanced memberships. Otherwise, no need to increase topic number.

```{r}
d <- 800
weight_shared <- 0.5
nlabel <- 5
```
```{r}
# 1. keep E(a) / (k0 + k1) constant
# 2. keep E(a) / 2k0 constant
# 3. keep E(a) / 2k1 constant
# comparison
for(k0 in seq(5, 31, by = 5)){

  k1 <- (d * 0.2) - nlabel * k0
  set.seed(0)
  
  C <- 10
  print(k0)
  for(b in c(1 / ((k0 + k1) * C), 1 / (2 * k0 * C), 1 / (2 * k1 * C))){
    nsample <- 200
    
    Gs <- matrix(NA, nrow = nsample, ncol = k0 + k1)
    for(i in 1:nsample){
      a <- rexp(n = 1, rate = b)
      rho1 <- gtools::rdirichlet(n = 1, alpha = rep(1, k0))
      rho2 <- gtools::rdirichlet(n = 1, alpha = rep(1, k1))  
      rho <- c((1 - weight_shared) * rho1, weight_shared * rho2)
      Gs[i,] <- gtools::rdirichlet(n = 1, alpha = a * rho)
    }
    
    p <- as.data.frame(Gs) %>%
      pivot_longer(cols = everything(), names_to = "t", values_to = "va") %>%
      ggplot(aes(x = t, y = va)) +
      geom_boxplot()
    print(p)
    p <- as.data.frame(Gs) %>%
      mutate(sample = 1:200) %>%
      pivot_longer(cols = !sample, names_to = "t", values_to = "va", names_prefix = "V") %>%
      arrange(t) %>%
      ggplot(aes(x = t, y = va, group = sample)) +
      geom_path()
    print(p)
  }
}
```
```{r}
# 1. keep E(a) / (k0 + k1) constant
# 2. keep E(a) / 2k0 constant
# 3. keep E(a) / 2k1 constant
# comparison
for(k0 in seq(5, 31, by = 5)){

  k1 <- (d * 0.2) - nlabel * k0
  set.seed(0)
  
  C <- 50
  print(k0)
  for(b in c(1 / ((k0 + k1) * C), 1 / (2 * k0 * C), 1 / (2 * k1 * C))){
    nsample <- 200
    
    Gs <- matrix(NA, nrow = nsample, ncol = k0 + k1)
    for(i in 1:nsample){
      a <- rexp(n = 1, rate = b)
      rho1 <- gtools::rdirichlet(n = 1, alpha = rep(1, k0))
      rho2 <- gtools::rdirichlet(n = 1, alpha = rep(1, k1))  
      rho <- c((1 - weight_shared) * rho1, weight_shared * rho2)
      Gs[i,] <- gtools::rdirichlet(n = 1, alpha = a * rho)
    }
    
    p <- as.data.frame(Gs) %>%
      pivot_longer(cols = everything(), names_to = "t", values_to = "va") %>%
      ggplot(aes(x = t, y = va)) +
      geom_boxplot()
    print(p)
    p <- as.data.frame(Gs) %>%
      mutate(sample = 1:200) %>%
      pivot_longer(cols = !sample, names_to = "t", values_to = "va", names_prefix = "V") %>%
      arrange(t) %>%
      ggplot(aes(x = t, y = va, group = sample)) +
      geom_path()
    print(p)
  }
}
```

We decide to choose $\frac{E(a)}{k_0+Kk_1}=C$ to make sure all topics are uniform enough.

### alpha
```{r}
d <- 800
weight_shared <- 0.5
nlabel <- 5
```
```{r}
k0 = 10
  k1 <- (d * 0.2) - nlabel * k0
  set.seed(0)
  nsample <- 200
  rhos <- matrix(NA, nrow = nsample, ncol = k0 + k1)
  rhos[1:nsample,1:k0] <- gtools::rdirichlet(n = nsample, alpha = rep(1, k0)) * (1 - weight_shared)
  rhos[1:nsample, (k0 + 1):(k0 + k1)] <- gtools::rdirichlet(n = nsample, alpha = rep(1, k1)) * weight_shared
  
  alpha0 <- c(rep(1 / k0, k0) * (1 - weight_shared), rep(1 / k1, k1) * weight_shared)
  rhos2 <- gtools::rdirichlet(n = nsample, alpha = alpha0)
apply(rhos, 2, mean)
apply(rhos2, 2, mean)
sum(apply(rhos, 2, mean)[1:k0])
sum(apply(rhos2, 2, mean)[1:k0])
```

1st idea: still learn with (1, 1, ...). Although this prior doesn't capture the weight, but it puts no preference.

### training setting: supervised

```{r}
k0 = 10
 k1 <- (d * 0.2) - nlabel * k0
  set.seed(0)
  
  C <- 50
  print(k0)
 
  nsample <- 200
  rhos <- matrix(NA, nrow = nsample, ncol = k0 + k1)
  rhos[1:nsample,1:k0] <- gtools::rdirichlet(n = nsample, alpha = rep(1, k0)) * (1 - weight_shared)
  rhos[1:nsample, (k0 + 1):k1] <- gtools::rdirichlet(n = nsample, alpha = rep(1, k0)) * weight_shared
  
 
      
```

```{r}
for(k0 in seq(5, 31, by = 5)){

  k1 <- (d * 0.2) - nlabel * k0
  set.seed(0)
  
  C <- 50
  print(k0)
 
  rho1 <- gtools::rdirichlet(n = 1, alpha = rep(1, k0))
      rho2 <- gtools::rdirichlet(n = 1, alpha = rep(1, k1))  
      rho <- c((1 - weight_shared) * rho1, weight_shared * rho2)

    
   
    
    p <- as.data.frame(Gs) %>%
      pivot_longer(cols = everything(), names_to = "t", values_to = "va") %>%
      ggplot(aes(x = t, y = va)) +
      geom_boxplot()
    print(p)
    p <- as.data.frame(Gs) %>%
      mutate(sample = 1:200) %>%
      pivot_longer(cols = !sample, names_to = "t", values_to = "va", names_prefix = "V") %>%
      arrange(t) %>%
      ggplot(aes(x = t, y = va, group = sample)) +
      geom_path()
    print(p)
  
}
```



### topics
```{r}
ntopic <- 30
K <- ntopic
d <- 3
alpha_Lambda <- 1
beta_Lambda <- 2
mu_Mu <- 0
sigma2_Mu <- 10
Lambda <- rgamma(n = ntopic * d, shape = alpha_Lambda, rate = beta_Lambda) %>%
  matrix(nrow = K)

Mu <- matrix(NA, nrow = K, ncol = d)
for (i in 1:K){
  for (j in 1:d){
    Mu[i, j] <- rnorm(1, mu_Mu, sqrt(sigma2_Mu / Lambda[i, j]))
  }
}
nsample <- 30
X <- matrix(NA, nrow = K * nsample, ncol = d)
for(i in 1:K){
  e <- i * nsample
  s <- e - nsample + 1
  for(j in 1:d){
    X[s:e, j] <- rnorm(n = nsample, mean = Mu[i, j], sd = sqrt(1/Lambda[i,j]))  
  }
}

as.data.frame(X) %>%
  mutate(topic = rep(1:K, each = nsample)) %>%
  ggplot(aes(x = V1, y = V2, color = factor(topic))) +
  geom_point()

############
sigma2_Mu <- 100
Mu <- matrix(NA, nrow = K, ncol = d)
for (i in 1:K){
  for (j in 1:d){
    Mu[i, j] <- rnorm(1, mu_Mu, sqrt(sigma2_Mu / Lambda[i, j]))
  }
}
nsample <- 30
X <- matrix(NA, nrow = K * nsample, ncol = d)
for(i in 1:K){
  e <- i * nsample
  s <- e - nsample + 1
  for(j in 1:d){
    X[s:e, j] <- rnorm(n = nsample, mean = Mu[i, j], sd = sqrt(1/Lambda[i,j]))  
  }
}

as.data.frame(X) %>%
  mutate(topic = rep(1:K, each = nsample)) %>%
  ggplot(aes(x = V1, y = V2, color = factor(topic))) +
  geom_point()
# Tau <- Mu * Lambda
# plot(Mu[,1], Mu[,2])
# as.data.frame(Mu) %>%
#   ggplot(aes(x = V1, y = V2, size =))
```
### choose topic prior
```{r}
alpha_Lambda <- 2
beta_Lambda <- 4
mu_Mu <- 0
sigma2_Mu <- 1
nd <- seq(0.01, 2, by = 0.05)
plot(nd, dgamma(nd, alpha_Lambda, rate = beta_Lambda))
nd <- seq(1, 20, by = 0.2)
plot(nd, dgamma(1/nd, alpha_Lambda, rate = beta_Lambda)/nd/nd)
```

```{r}
source("../../R/shared.R")
source("../../R/Normal.R")
d <- 800
weight_shared <- 0.5
nlabel <- 5
k0 = 10
 k1 <- (d * 0.2) - nlabel * k0
 
 seed <- 2022
 
 dg <- k0 + k1
K <- nlabel * k0 + k1
ntopic <- K

Ts <- T_generator(k0, k1, nlabel)

set.seed(seed)

# sample parameters
b <- 20 / (k0 + k1)
a <- rexp(n = 1, rate = b)

rho1 <- gtools::rdirichlet(n = 1, alpha = rep(1, k0))
rho2 <- gtools::rdirichlet(n = 1, alpha = rep(1, k1))
rho <- c((1 - weight_shared) * rho1, weight_shared * rho2)

Lambda <- rgamma(n = ntopic * d, shape = alpha_Lambda, rate = beta_Lambda) %>%
  matrix(nrow = K)

Mu <- matrix(NA, nrow = K, ncol = d)
for (i in 1:K){
  for (j in 1:d){
    Mu[i, j] <- rnorm(1, mu_Mu, sqrt(sigma2_Mu / Lambda[i, j]))
  }
}
Mu <- Mu * sqrt(10)
Tau <- Mu * Lambda

dat <- document_generator_Normal(a, rho, Ts, Lambda, Tau, 200, NULL, seed)

```


```{r}
d1 <- 1
d2 <- 2
dat$X[,c(d1, d2)] %>% as.data.frame() %>%
  mutate(label = factor(dat$Y)) %>%
  ggplot(aes(x = V1, y = V2, color = label)) +
  geom_point()

prediction <- SPM_predicting_Normal(dat$X[1:100,], Lambda, Mu, a, rho, Ts, 500, seed) 

accuracy <- sum(prediction[["labels"]] == dat[["Y"]][1:100])
accuracy # 0.94 for 1, 0.96 for 10
```



